---
title: "Lab report knapsack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab report knapsack}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Knapsack)
```

## The knapsack problem

The Knapsack package contains two different ways of solving the so called knapsack problem. The knapsack problem is a discrete optimization problem where we have a knapsack that can take a limited weight $W$ and we want to fill this knapsack with a number of items $i = 1,\cdots, n$, each with weight $w_i$ and a value $v_i$. The goal is to find the knapsack with the largest value of the elements added to the knapsack.

The data used in this assignemt is generated in the following way

```{r}
set.seed(42)
n <- 2000
knapsack_objects <- data.frame(w = sample(1: 4000, size = n, replace = TRUE),v = runif(n = n, 0, 10000))
```

## `brute_force_knapsack()`

The only solution that is guaranteed to give a correct answer in all situations for the knapsack problem
is using brute-force search, i.e. going through all possible alternatives and return the maximum value
found.This is done by the function `brute_force_knapsack(x, W)`.^[Assignment 4 by Ellinor Fackle Fornius & Jessica Franzén.]

The function takes two arguments, a `data.frame x` with two variables, `v`and `w`, and `W`, the knapsack size. It then returns the maximum knapsack value and which elements are used to acheive that value.


## How much time does it take to run the algorithm for $n = 16$ objects?

This will depend on W. I included a line in the beginning of the function that filters out all items $i$ where $w_i > W$, since these wouldn't fit in the knapsack anyways. Given $n$ objects, there are $2^n$ possible combinations of objects that are gone through using this approach. The smaller $W$ is, then reasonably more objects will be excluded. For each item that is excluded from the start, the number of possible combinations decrease by half. 

```{r, echo=TRUE, eval=TRUE}
system.time(brute_force_knapsack(x = knapsack_objects [1: 16,],W = 2000))
```

```{r, echo=TRUE, eval=TRUE}
system.time(brute_force_knapsack(x = knapsack_objects [1: 16,],W = 4000))
```

## `greedy_knapsack()`

This function gives an approximation for the problem by using George Dantzigs greedy approximation algorithm.,
The items are sorted in decreasing order of value per unit of weight, $v_1/w_1 \geq \cdots \geq v_n/w_n$. The function then constructs a solution $S_1$ by packing items greedily until the maximum weight limit is reached, i.e. $S_1 = \{1,\cdots, k\}$ where $k = \max_{1\leq k'\leq n} \sum_{i=1}^k w_i \leq W$. A second solution $S_2=\{k+1\}$, containing the first element that did not fit, is then constructed.Since $S_1 \cup S_2$ provides an upper bound for the LP relaxation of the problem, one of them must have value at least $50\%$ of the maximum value of items that fit into the sack. Thus, the function returns the solution with the greatest value.^[Wikipedia [via](https://en.wikipedia.org/wiki/Knapsack_problem#Greedy_approximation_algorithm)]


## How much time does it take to run the algorithm for $n = 1000000$ objects?

```{r}
set.seed(42)
n <- 1000000
knapsack_objects <- data.frame(w = sample(1: 4000, size = n, replace = TRUE),v = runif(n = n, 0, 10000))
system.time(greedy_knapsack(x = knapsack_objects,W = 2000))
```


